Applying Discontinuous Petrov-Galerkin Methods to Fluid Problems

### 2012-03-23 docs/presentations/nate/Proposal/paper.tex
Updated Stokes write-up.
git-svn-id: https://svn.ices.utexas.edu/repos/camellia@387 36274a68-8941-4c93-80ef-affaed5139da

\begin{align}
- \mu \Delta \NVRvect{u} + \NVRgrad p &= \NVRvect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst1} \\
\NVRdiv \NVRvect{u} &= 0 & \text{ in } \Omega, \\
\NVRvect{u} &= \NVRvect{g_{D}} & \text{ on } \partial\Omega,\label{NVR:eq:stronglast1}
\end{align}
where $\Omega \subset \mathbb{R}^{2}$, $\mu$ is viscosity, $\NVRtensor{\epsilon}=\NVRgrad^{\text{sym}} \NVRvect{u}$ is strain, $p$ is pressure, $u$ velocity, and $\NVRvect{f}$ a vector forcing function.

For DPG, we require a first-order system.  We introduce $\NVRtensor{\sigma}=\mu \NVRgrad \NVRvect{u}$:
\begin{align}
\frac{1}{\mu} \NVRtensor{\sigma} - \NVRgrad \NVRvect{u} &= 0 \\
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \NVRvect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst2} \\
\NVRdiv \NVRvect{u} &= 0 & \text{ in } \Omega, \\
\NVRvect{u} &= \NVRvect{g_{D}} & \text{ on } \partial\Omega,\label{NVR:eq:stronglast2}
\end{align}

@@

\section{DPG Method}\label{NVR:sec:DPGintro}
Here, we sketch some of the main features of the DPG method.  For details, we refer the reader to a series of papers by Demkowicz et al., in particular the second ICES Report \cite{DPG2}, from which most of this section is derived.  We begin with theoretical definitions and results, and then describe the approach to practical realization.  Consider the abstract variational boundary-value problem:
\begin{align}
\text{Find } u \in U : b(u,v) = l(v) \quad \forall v \in V.\label{NVR:eq:avbvp}
\end{align}
We take $U$ and $V$ to be real Hilbert spaces.  We assume $b(\cdot,\cdot)$ is continuous, i.e.\
\begin{align}
|b(u,v)| \leq M \NVRnorm{u}_{U} \NVRnorm{v}_{V},
\end{align}
for some real $M$. We assume also that $b(\cdot,\cdot)$ is weakly coercive, that is
\begin{align}
\inf_{\NVRnorm{u}_{U}=1} \sup_{\NVRnorm{v}_{V}=1} b(u,v) > \gamma,
\end{align}
for some $\gamma > 0$.  If we additionally assume that
\begin{align}
\{v \in V : b(u,v)=0 \quad \forall u \in U\} = \{0\},
\end{align}
then it is well known that the problem \eqref{NVR:eq:avbvp} has a unique solution provided that $l \in V'$, the dual of $V$.

\subsection{Energy Norm}
We define an alternate norm, called the \emph{energy norm}, on the trial space $U$ by
\begin{equation}\label{NVR:eq:energyNorm}
\NVRnorm{u}_{E} \NVReqdef \sup_{\NVRnorm{v}_{V}=1} b(u,v).
\end{equation}
This norm is the one in which the optimality is guaranteed by the selection of optimal test functions. It is an equivalent norm to the standard norm on $U$, i.e.\
\begin{equation}
\gamma \NVRnorm{u}_{U} \leq \NVRnorm{u}_{E} \leq M \NVRnorm{u}_{U} \quad \forall u \in U.
\end{equation}

\subsection{Optimal Test Functions}
We are now prepared to give a definition of the optimal test functions.  Define a map $T:U \rightarrow V$ from the trial space to the test space by: For $u \in U$, define $Tu$, the \emph{optimal test function} corresponding to $u$, as the unique solution to 
\begin{align*}
(Tu,v)_{V} = b(u,v) \quad \forall v \in V.
\end{align*}
By the Riesz representation theorem, $T$ is well-defined.  Note that 
\begin{align*}
\NVRnorm{u}_{E} = \sup_{\NVRnorm{v}_{V}=1} b(u,v) = \sup_{\NVRnorm{v}_{V}=1} (Tu,v)_{V} = \frac{1}{\NVRnorm{Tu}_{V}} (Tu, Tu)_{V} = \NVRnorm{Tu}_{V}.
\end{align*}
Thus the energy norm is generated by the inner product on $V$, i.e.\ 
\begin{equation}
(u,u)_{E} \NVReqdef (Tu,Tu)_{V}.\label{NVR:eq:ipEqualsEnergy}
\end{equation}
In practice, we approximate $T$ by a discrete operator $T_{n}$, described in Section \ref{NVR:sec:practicalRealization}.
\subsection{Optimal Test Space for $U_{n}$}
Take a finite-dimensional trial space $U_{n} \subset U$.  Define the \emph{optimal test space} for $U_{n}$ as $V_{n} = \text{span} \{Te_{j} : j=1,\cdots,n\}$, where the $e_{j}$ form a basis for $U_{n}$.

Solve the discrete problem
\begin{equation}
\text{Find } u_{n} \in U_{n} : b(u_{n}, v) = l(v) \ \forall v \in V_{n}. \label{NVR:eq:discreteAVBVP}
\end{equation}
Then the error is the best approximation error in the energy norm,
\begin{align}
\NVRnorm{u-u_{n}}_{E} = \inf_{w_{n} \in U_{n}} \NVRnorm{u-w_{n}}_{E},\label{NVR:eq:optimalityEqn}
\end{align}
and this is the sense in which the test space is \emph{optimal}.

@@

\section{Two Standard Choices for Test Space Norm}\label{NVR:sec:testNorms}
In this section, we describe two commonly used norms on the test space --- recall that the choice of test space norm will determine the energy norm, and thus the sense in which test functions are optimal.  Our implementation makes usage of these norms simple.
\subsection{Mathematician's Norm}
In general, test functions in a variational form each belong to some functional space on which a norm is already defined.  Therefore, a mathematically natural choice for a norm on the test space will be the Euclidean combination of these norms.  In the case of our Poisson formulation, $\NVRvect{q} \in \NVRHdiv$ and $v \in \NVRHgrad$, so define
\begin{align*}
\NVRnorm{(\NVRvect{q},v)}_{V}^{2}  \NVReqdef \int_{\Omega} &\left((\NVRdiv \NVRvect{q})^{2} + \NVRvect{q} \cdot \NVRvect{q} + \NVRgrad v \cdot \NVRgrad v + v^{2}\right)\\
\end{align*}
as the \emph{mathematician's} norm on the test space.

\subsection{Optimal and Quasi-Optimal Test Norms}
We know from the analysis that DPG delivers the best solution in the \emph{energy norm}, and that the choice of norm on the test space determines the energy norm.  We can then ask whether it is possible to select the test space norm in such a way that the energy norm is exactly the norm of interest ($L^{2}$, in our case) --- and the answer is yes; this is given in the abstract setting by 
\begin{align*}
\NVRnorm{v}_{V}\NVReqdef \sup_{\NVRnorm{u}_{U}=1} b(u,v),
\end{align*}
where $\NVRnorm{\cdot}_{U}$ is the norm of interest.  $\NVRnorm{\cdot}_{V}$ is then called the \emph{optimal test norm}.  To determine the optimal test norm, we can collect terms by trial space variable, and then take the supremum using the Cauchy-Schwarz inequality.  When we do so, we will have some element boundary terms arising from the fluxes and traces.  To make the test norm \emph{localizable}, we then replace these terms with $L^{2}$ terms on the element interior, often with some weight $\beta$.  The resultant norm we call the \emph{quasi-optimal test norm}.

@@

\begin{align}
- 2 \mu \NVRdiv \NVRtensor{\epsilon} + \NVRgrad p &= \NVRvect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst} \\
\NVRdiv \NVRvect{u} &= 0 & \text{ in } \Omega, \\
\NVRvect{u} &= \NVRvect{g_{D}} & \text{ on } \partial\Omega,\label{NVR:eq:stronglast}
\end{align}
where $\Omega \subset \mathbb{R}^{2}$, $\mu$ is viscosity, $\NVRtensor{\epsilon}=\NVRgrad^{\text{sym}} \NVRvect{u}$ is strain, $p$ is pressure, $u$ velocity, and $\NVRvect{f}$ a vector forcing function.

We introduce stress $\sigma$ and vorticity $\omega$ by
\begin{align*}
\NVRtensor{\sigma} &= 2 \mu \NVRtensor{\epsilon} - p\NVRtensor{I}\\
\NVRtensor{\omega} &= \frac{1}{2}(\NVRgrad \NVRvect{u} - \NVRgrad \NVRvect{u}^{T})
\end{align*}
so that equation \eqref{NVR:eq:strongfirst} becomes simply $-\NVRdiv \NVRtensor{\sigma} = \NVRvect{f}$.
We also have
\begin{align*}
\NVRtensor{\epsilon} &= \frac{1}{2 \mu}(\NVRtensor{\sigma} + p\NVRtensor{I}).
\end{align*}
Since $\NVRtensor{\epsilon} = \NVRgrad^{\text{sym}} \NVRvect{u} = \NVRgrad \NVRvect{u} - \NVRtensor{\omega}$, the entire system is

\begin{align*}
\frac{1}{2 \mu}(\NVRtensor{\sigma} + p\NVRtensor{I}) - \NVRgrad \NVRvect{u} + \NVRtensor{\omega} &= \NVRtensor{0}  & \text{ in } \Omega, \\
- \NVRdiv \NVRtensor{\sigma} &= \NVRvect{f} & \text{ in } \Omega, \\
\NVRdiv \NVRvect{u} &= 0 & \text{ in } \Omega, \\
\NVRvect{u} &= \NVRvect{g_{D}} & \text{ on } \partial\Omega.
\end{align*}
Note that the antisymmetric part of the first equation recovers the definition of $\NVRtensor{\omega}$, so that it need not enter the system separately.
Define scalar $\omega = \omega_{21} = \frac{1}{2}(u_{1,2} - u_{2,1})$.  Our strong formulation is

\begin{align*}
\frac{1}{2 \mu}{\sigma_{11} + p \choose \sigma_{21}} - \NVRgrad u_{1} + {0 \choose \omega} &= \NVRvect{0} & \text{ in } \Omega, \\
\frac{1}{2 \mu}{\sigma_{12} \choose \sigma_{22} + p} - \NVRgrad u_{2} - {\omega \choose 0} &= \NVRvect{0}  & \text{ in } \Omega, \\
-\NVRdiv {\sigma_{11} \choose \sigma_{21}} &= f_{1} & \text{ in } \Omega, \\
-\NVRdiv {\sigma_{12} \choose \sigma_{22}} &= f_{2} & \text{ in } \Omega, \\
\NVRdiv \NVRvect{u} &= 0 & \text{ in } \Omega, \\\
\NVRvect{u} &= \NVRvect{g_{D}} & \text{ on } \partial\Omega.
\end{align*}
Multiplying the first two equations by vector test functions $\NVRvect{q}_{i}$ and the following three by scalar test functions $v_{i}$, and integrating by parts over an element $K$, we obtain

>>>>

\begin{align}
- \mu \Delta \vect{u} + \NVRgrad p &= \vect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst1} \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\
\vect{u} &= \vect{g_{D}} & \text{ on } \partial\Omega,\label{NVR:eq:stronglast1}
\end{align}
where $\Omega \subset \mathbb{R}^{2}$, $\mu$ is viscosity, $p$ is pressure, $\vect{u}$ velocity, and $\vect{f}$ a vector forcing function.

For DPG, we require a first-order system.  We introduce $\NVRtensor{\sigma}= \NVRgrad \vect{u}$:
\begin{align}
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \vect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst2} \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\
\NVRtensor{\sigma} - \NVRgrad \vect{u} &= 0 & \text{ in } \Omega,\\
\vect{u} &= \vect{g_{D}} & \text{ on } \partial\Omega.\label{NVR:eq:stronglast2}
\end{align}

Considering the above as a linear operator $A$ operating on a group variable $\tilde{u} = (\vect{u},p,\NVRtensor{\sigma})$, to apply the abstract theory presented above we need to establish that $A$ and $A^{*}$ are injective, and either that $||A \tilde{u}|| \geq \gamma ||\tilde{u}||$ or that $||A^{*} \tilde{v}|| \geq \gamma ||\tilde{v}||$.  Since the latter two conditions imply the injectivity of the respective operators, it will suffice to prove these.\\

\subsection{$||A \tilde{u}|| \geq \gamma ||\tilde{u}||$}
Adding right-hand sides $g$ and $\NVRtensor{h}$

\begin{align}
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \vect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst3} \\
\NVRdiv \vect{u} &= g & \text{ in } \Omega, \\
\NVRtensor{\sigma} - \NVRgrad \vect{u} &= \NVRtensor{h} & \text{ in } \Omega, \label{NVR:eq:strongsecondtolast3} \\
\vect{u} &= \vect{g_{D}} & \text{ on } \partial\Omega,\label{NVR:eq:stronglast3}
\end{align}
we have that $A \tilde{u} = (\vect{f},g,\NVRtensor{h})$.  If we can establish bounds for each of the solution variables in terms of norms on $\vect{f},g,$ and $\NVRtensor{h}$, then we will have the required lower bound $|| A \tilde{u} || \geq \gamma || \tilde{u} ||$.  For the time being we assume that $\vect{g_{D}} = 0$.\\

Note that, by linearity of $A$, it suffices to consider cases in which only one of $\vect{f},g,$ and $\NVRtensor{h}$ is non-zero.\\

\subsubsection{$\vect{f} \neq \vect{0}$}
In this case, we have exactly the system (\ref{NVR:eq:strongfirst2})-(\ref{NVR:eq:stronglast2}), which reduces (in a distributional sense) to (\ref{NVR:eq:strongfirst1})-(\ref{NVR:eq:stronglast1}).  Testing the first equation with the velocity $u \in H^{1}_{0}(\Omega)$, we have
\begin{align*}
- (\mu \Delta \vect{u}, \vect{u})_{\Omega} + (\NVRgrad p, \vect{u})_{\Omega} &= (\vect{f},\vect{u})_{\Omega}.\\
\end{align*}
Integrating by parts, we obtain:
\begin{align*}
(\mu \NVRgrad \vect{u}, \NVRgrad \vect{u})_{\Omega} - \langle \NVRgrad \vect{u} \cdot \vect{n}, \vect{u} \rangle_{\partial \Omega} - (p, \NVRdiv u)_{\Omega} + \langle p, \vect{u} \cdot \vect{n} \rangle_{\partial \Omega} &= (\vect{f},\vect{u})_{\Omega}.
\end{align*}

Noting that $\vect{u} = 0$ on $\partial \Omega$ and that $\NVRdiv \vect{u} = 0 $ in $\Omega$, this reduces to:

\begin{align*}
\mu (\NVRgrad \vect{u}, \NVRgrad \vect{u})_{\Omega} = ||\NVRgrad \vect{u}||^{2}_{\Omega} = || \NVRtensor{ \sigma }||^{2}_{\Omega} &= (\vect{f},\vect{u})_{\Omega} \\
&\leq || f ||_{\Omega} || \vect{u} ||_{\Omega}\\
&\leq C_{P} || f ||_{\Omega} || \NVRgrad \vect{u} ||_{\Omega},
\end{align*}
where $C_{P}$ is the Poincare constant.  Thus $ || \NVRtensor{ \sigma }||_{\Omega} \leq \frac{C_{P}}{\mu} || f ||_{\Omega}$, and $|| \vect{u} || \leq \frac{C_{P}^{2}}{\mu} || f ||_{\Omega}$.

To bound the pressure $p$, we require a result by Ladyzhenskaya for $p \in L^{2}_{0}$:
\begin{align*}
\sup_{\vect{v} \in H^{1}(\Omega)} \frac{(p, \NVRdiv \vect{v})_{\Omega}}{\norm{ \vect{v} }_{H^{1}(\Omega)} } \geq \beta \norm{p}_{\Omega}
\end{align*}
for some constant $\beta > 0$.  Testing equation (\ref{NVR:eq:strongfirst1}) with $\vect{v} \in \vect{H}^{1}_{0}(\Omega)$, we have
\begin{align*}
\NVRgrad p \cdot \vect{v} = \vect{f} \cdot \vect{v} + \mu \Delta \vect{u} \cdot \vect{v}.
\end{align*}

Integrating by parts, dividing by $\norm{ \vect{v} }_{H^{1}(\Omega)}$ and taking the supremum:
\begin{align*}
\sup - \frac{(p, \NVRdiv \vect{v})}{\norm{ \vect{v} }_{H^{1}(\Omega)}} &= \sup \left\{\frac{(\vect{f}, \vect{v})}{\norm{ \vect{v} }_{H^{1}(\Omega)}} - \frac{\mu (\NVRgrad \vect{u}, \NVRgrad \vect{v})}{\norm{ \vect{v} }_{H^{1}(\Omega)}}\right\} \\
&\leq \frac{\norm{\vect{f}} \norm{\vect{v}}_{L^{2}}}{\norm{ \vect{v} }_{H^{1}(\Omega)}} + \mu \norm{\NVRgrad \vect{u}} \leq \norm{\vect{f} } + \mu \frac{C_{P}}{\mu} \norm{\vect{f}} = (1 + C_{P}) \norm{\vect{f}}.
\end{align*}
By the Ladyzhenskaya result, we then have $\norm{p} \leq \frac{1 + C_{P}}{\beta} \norm{\vect{f}}$, bounding $p$, as required.

\subsubsection{$ g \neq 0$} In this case, we have
\begin{align}
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \vect{0} & \text{ in } \Omega, \label{NVR:eq:gstrongfirst} \\
\NVRdiv \vect{u} &= g & \text{ in } \Omega, \\
\NVRtensor{\sigma} - \NVRgrad \vect{u} &= 0 & \text{ in } \Omega,\\
\vect{u} &= \vect{0} & \text{ on } \partial\Omega.\label{NVR:eq:gstronglast}
\end{align}

We must also assume compatibility between $g$ and the boundary condition on $\vect{u}$; that is, that $g$ has zero average on $\Omega$:
\begin{align*}
\int_{\Omega} g = \int_{\partial \Omega} \vect{u} \cdot \vect{n} = 0.
\end{align*}

Now, by surjectivity of $\NVRdiv$, 
\begin{align*}
\forall g \in L^{2} \quad \exists \vect{u}_{0} : \NVRdiv \vect{u}_{0} = g.
\end{align*}

By the Lax-Milgram theorem, we have $\exists ! \phi \in H^{1}_{0} : \vect{u}_{0} = \NVRgrad \phi$, so that $\Delta \phi = g$ and, assuming that $\Omega$ is convex, then by the Elliptic Regularity theorem (see \cite[p. 214]{Folland} for a proof), $\phi \in H^{2}$ and:
\begin{align*}
\norm{\phi}_{H^{2}} &\leq C \norm{g}\\ % SEE ARBOGAST/BONA, p. 223 (Elliptic Regularity, section 8.5.4, Theorem 8.13)---references [GT] or [Fo] for a proof
\norm{\NVRgrad \phi}_{H^{1}} &\leq C \norm{g},
\end{align*}
for some $C > 0$ independent of $g$ and $\phi$.  That is, in fact $\vect{u}_{0} \in H^{1}$ and $\norm{ \vect{u}_{0} }_{H^{1}} \leq C \norm{g}$. \\

Now, setting $\vect{w} = \vect{u} - \vect{u}_{0}$, we have $\NVRdiv \vect{w} = 0$.  Testing equation (\ref{NVR:eq:strongfirst1}) with $\vect{w}$ and integrating by parts, we obtain:
\begin{align*}
\mu (\NVRgrad \vect{u},\NVRgrad \vect{w})_{\Omega} &= 0\\
\mu (\NVRgrad \vect{w} + \NVRgrad \vect{u}_{0},\NVRgrad \vect{w})_{\Omega} &= 0\\
\implies \mu \norm{\NVRgrad \vect{w}}^{2} &= -(\NVRgrad \vect{u}_{0}, \NVRgrad \vect{w}) \\
&\leq \norm{ \NVRgrad \vect{u}_{0} } \norm{ \NVRgrad \vect{w} } \\
&\leq C \norm{g} \norm{ \NVRgrad \vect{w} },
\end{align*}
so that $\norm{ \NVRgrad \vect{w} } \leq  \frac{C}{\mu} \norm{ g }$, and therefore as before we obtain a bound on $\norm{\NVRtensor{\sigma}} = \norm{\NVRgrad \vect{u}} \leq \left( 1 + \frac{1}{\mu} \right) C \norm{g}$, and again by means of the Poincare inequality we may bound $\norm{\vect{u}} \leq C_{P} \norm{\NVRgrad \vect{u}} \leq \frac{C C_{P}}{\mu} \norm{ g }$ as well.\\

The bound on $p$ is established exactly as before, except that now $\vect{f} = \vect{0}$, so that we obtain the bound $\norm{p} \leq \frac{\mu}{\beta} \norm{\NVRgrad \vect{u}}$.

\subsubsection{$ \NVRtensor{h} \neq \NVRtensor{0}$}  In this case, we have
\begin{align}
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \vect{0} & \text{ in } \Omega, \label{NVR:eq:hstrongfirst} \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\
\NVRtensor{\sigma} - \NVRgrad \vect{u} &= \NVRtensor{h} & \text{ in } \Omega, \\
\vect{u} &= \vect{0} & \text{ on } \partial\Omega.\label{NVR:eq:hstronglast}
\end{align}

Then, $\NVRtensor{\sigma} = \NVRgrad \vect{u} + \NVRtensor{h}$, and we have:
\begin{align*}
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= - \mu \Delta \vect{u} - \mu \NVRdiv \NVRtensor{h} + \NVRgrad p = \vect{0} \\
- \mu \Delta \vect{u} + \NVRgrad p &= - \mu  \NVRdiv \NVRtensor{h}
\end{align*}

in a distributional sense.  Testing with $\vect{u}$, and integrating by parts, again the pressure term vanishes because $\NVRdiv \vect{u} = 0$, so that much as before, we obtain:

\begin{align*}
\mu (\NVRgrad \vect{u}, \NVRgrad \vect{u})_{\Omega} &= \langle \NVRdiv \NVRtensor{h}, \vect{u} \rangle_{\partial \Omega}\\
&= -(\NVRtensor{h}, \NVRgrad \vect{u})_{\Omega} \leq \norm{\NVRtensor{h}}_{L^{2}(\Omega)} \norm{ \NVRgrad \vect{u} }_{L^{2}(\Omega)}.
\end{align*}

So $\norm{\NVRgrad \vect{u}} \leq \frac{1}{\mu} \norm{\NVRtensor{h}}$, and the bounds $\norm{\vect{u} } \leq \frac{C_{P}}{\mu} \norm{\NVRtensor{h}}$ and $\norm{p} \leq \frac{\mu}{\beta}\left( \norm{ \NVRgrad \vect{u} } + \norm{\NVRtensor{h}} \right)$ can be established in a similar fashion as above.

%% ADJOINT %%
\subsection{$||A^{*} \tilde{v}|| \geq \gamma ||\tilde{v}||$}  If we consider $(A \tilde{u}, \tilde{v})$, where $\tilde{v} = (\NVRtensor{\tau},\vect{q}, v)$, and integrate by parts (ignoring boundary terms for now), we can readily see that the adjoint equation $A^{*} \tilde{v} = \vect{F}$ takes the form:
\begin{align}
\NVRdiv \NVRtensor{\tau} - \NVRgrad v &= \vect{f^{*}} & \text{ in } \Omega, \\
- \NVRdiv \vect{q} &= g^{*} & \text{ in } \Omega, \\
\NVRtensor{\tau} + \mu \NVRgrad \vect{q} &= \NVRtensor{h^{*}} & \text{ in } \Omega.
\end{align}

If we then substitute $\NVRtensor{\tau} = - \mu \NVRtensor{\sigma}$ and $v = -p$, we have:
\begin{align}
- \mu \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \vect{f}^{*} & \text{ in } \Omega, \\
\NVRdiv \vect{q} &= -g^{*} & \text{ in } \Omega, \\
\NVRtensor{\sigma} - \NVRgrad \vect{q} &= - \frac{1}{\mu} \NVRtensor{h}^{*} & \text{ in } \Omega,
\end{align}

so that the adjoint $A^{*}$ takes exactly the same form as the original operator $A$ (compare equations (\ref{NVR:eq:strongfirst3}) - (\ref{NVR:eq:strongsecondtolast3})).  Thus our argument for $||A \tilde{u}|| \geq \gamma ||\tilde{u}||$ will also work to show that $||A^{*} \tilde{v}|| \geq \gamma ||\tilde{v}||$.

@@

\section{DPG Method}\label{NVR:sec:DPGintro}
Here, we sketch some of the main features of the DPG method.  For details, we refer the reader to a series of papers by Demkowicz et al., in particular the second ICES Report \cite{DPG2}, from which most of this section is derived.  We begin with theoretical definitions and results, and then describe the approach to practical realization.  Consider the abstract variational boundary-value problem:
\begin{align}
\text{Find } u \in U : b(u,v) = l(v) \quad \forall v \in V.\label{NVR:eq:avbvp}
\end{align}
We take $U$ and $V$ to be real Hilbert spaces.  We assume $b(\cdot,\cdot)$ is continuous, i.e.\
\begin{align}
|b(u,v)| \leq M \norm{u}_{U} \norm{v}_{V},
\end{align}
for some real $M$. We assume also that $b(\cdot,\cdot)$ is weakly coercive, that is
\begin{align}
\inf_{\norm{u}_{U}=1} \sup_{\norm{v}_{V}=1} b(u,v) > \gamma,
\end{align}
for some $\gamma > 0$.  If we additionally assume that
\begin{align}
\{v \in V : b(u,v)=0 \quad \forall u \in U\} = \{0\},
\end{align}
then it is well known that the problem \eqref{NVR:eq:avbvp} has a unique solution provided that $l \in V'$, the dual of $V$.

\subsection{Energy Norm}
We define an alternate norm, called the \emph{energy norm}, on the trial space $U$ by
\begin{equation}\label{NVR:eq:energyNorm}
\norm{u}_{E} \NVReqdef \sup_{\norm{v}_{V}=1} b(u,v).
\end{equation}
This norm is the one in which the optimality is guaranteed by the selection of optimal test functions. It is an equivalent norm to the standard norm on $U$, i.e.\
\begin{equation}
\gamma \norm{u}_{U} \leq \norm{u}_{E} \leq M \norm{u}_{U} \quad \forall u \in U.
\end{equation}

\subsection{Optimal Test Functions}
We are now prepared to give a definition of the optimal test functions.  Define a map $T:U \rightarrow V$ from the trial space to the test space by: For $u \in U$, define $Tu$, the \emph{optimal test function} corresponding to $u$, as the unique solution to 
\begin{align*}
(Tu,v)_{V} = b(u,v) \quad \forall v \in V.
\end{align*}
By the Riesz representation theorem, $T$ is well-defined.  Note that 
\begin{align*}
\norm{u}_{E} = \sup_{\norm{v}_{V}=1} b(u,v) = \sup_{\norm{v}_{V}=1} (Tu,v)_{V} = \frac{1}{\norm{Tu}_{V}} (Tu, Tu)_{V} = \norm{Tu}_{V}.
\end{align*}
Thus the energy norm is generated by the inner product on $V$, i.e.\ 
\begin{equation}
(u,u)_{E} \NVReqdef (Tu,Tu)_{V}.\label{NVR:eq:ipEqualsEnergy}
\end{equation}
In practice, we approximate $T$ by a discrete operator $T_{n}$, described in Section \ref{NVR:sec:practicalRealization}.
\subsection{Optimal Test Space for $U_{n}$}
Take a finite-dimensional trial space $U_{n} \subset U$.  Define the \emph{optimal test space} for $U_{n}$ as $V_{n} = \text{span} \{Te_{j} : j=1,\cdots,n\}$, where the $e_{j}$ form a basis for $U_{n}$.

Solve the discrete problem
\begin{equation}
\text{Find } u_{n} \in U_{n} : b(u_{n}, v) = l(v) \ \forall v \in V_{n}. \label{NVR:eq:discreteAVBVP}
\end{equation}
Then the error is the best approximation error in the energy norm,
\begin{align}
\norm{u-u_{n}}_{E} = \inf_{w_{n} \in U_{n}} \norm{u-w_{n}}_{E},\label{NVR:eq:optimalityEqn}
\end{align}
and this is the sense in which the test space is \emph{optimal}.

@@

\section{Two Standard Choices for Test Space Norm}\label{NVR:sec:testNorms}
In this section, we describe two commonly used norms on the test space --- recall that the choice of test space norm will determine the energy norm, and thus the sense in which test functions are optimal.  Our implementation makes usage of these norms simple.
\subsection{Mathematician's Norm}
In general, test functions in a variational form each belong to some functional space on which a norm is already defined.  Therefore, a mathematically natural choice for a norm on the test space will be the Euclidean combination of these norms.  In the case of our Poisson formulation, $\vect{q} \in \NVRHdiv$ and $v \in \NVRHgrad$, so define
\begin{align*}
\norm{(\vect{q},v)}_{V}^{2}  \NVReqdef \int_{\Omega} &\left((\NVRdiv \vect{q})^{2} + \vect{q} \cdot \vect{q} + \NVRgrad v \cdot \NVRgrad v + v^{2}\right)\\
\end{align*}
as the \emph{mathematician's} norm on the test space.

\subsection{Optimal and Quasi-Optimal Test Norms}
We know from the analysis that DPG delivers the best solution in the \emph{energy norm}, and that the choice of norm on the test space determines the energy norm.  We can then ask whether it is possible to select the test space norm in such a way that the energy norm is exactly the norm of interest ($L^{2}$, in our case) --- and the answer is yes; this is given in the abstract setting by 
\begin{align*}
\norm{v}_{V}\NVReqdef \sup_{\norm{u}_{U}=1} b(u,v),
\end{align*}
where $\norm{\cdot}_{U}$ is the norm of interest.  $\norm{\cdot}_{V}$ is then called the \emph{optimal test norm}.  To determine the optimal test norm, we can collect terms by trial space variable, and then take the supremum using the Cauchy-Schwarz inequality.  When we do so, we will have some element boundary terms arising from the fluxes and traces.  To make the test norm \emph{localizable}, we then replace these terms with $L^{2}$ terms on the element interior, often with some weight $\beta$.  The resultant norm we call the \emph{quasi-optimal test norm}.

@@

\begin{align}
- 2 \mu \NVRdiv \NVRtensor{\epsilon} + \NVRgrad p &= \vect{f} & \text{ in } \Omega, \label{NVR:eq:strongfirst} \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\
\vect{u} &= \vect{g_{D}} & \text{ on } \partial\Omega,\label{NVR:eq:stronglast}
\end{align}
where $\Omega \subset \mathbb{R}^{2}$, $\mu$ is viscosity, $\NVRtensor{\epsilon}=\NVRgrad^{\text{sym}} \vect{u}$ is strain, $p$ is pressure, $u$ velocity, and $\vect{f}$ a vector forcing function.

We introduce stress $\sigma$ and vorticity $\omega$ by
\begin{align*}
\NVRtensor{\sigma} &= 2 \mu \NVRtensor{\epsilon} - p\NVRtensor{I}\\
\NVRtensor{\omega} &= \frac{1}{2}(\NVRgrad \vect{u} - \NVRgrad \vect{u}^{T})
\end{align*}
so that equation \eqref{NVR:eq:strongfirst} becomes simply $-\NVRdiv \NVRtensor{\sigma} = \vect{f}$.
We also have
\begin{align*}
\NVRtensor{\epsilon} &= \frac{1}{2 \mu}(\NVRtensor{\sigma} + p\NVRtensor{I}).
\end{align*}
Since $\NVRtensor{\epsilon} = \NVRgrad^{\text{sym}} \vect{u} = \NVRgrad \vect{u} - \NVRtensor{\omega}$, the entire system is

\begin{align*}
\frac{1}{2 \mu}(\NVRtensor{\sigma} + p\NVRtensor{I}) - \NVRgrad \vect{u} + \NVRtensor{\omega} &= \NVRtensor{0}  & \text{ in } \Omega, \\
- \NVRdiv \NVRtensor{\sigma} &= \vect{f} & \text{ in } \Omega, \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\
\vect{u} &= \vect{g_{D}} & \text{ on } \partial\Omega.
\end{align*}
Note that the antisymmetric part of the first equation recovers the definition of $\NVRtensor{\omega}$, so that it need not enter the system separately.
Define scalar $\omega = \omega_{21} = \frac{1}{2}(u_{1,2} - u_{2,1})$.  Our strong formulation is

\begin{align*}
\frac{1}{2 \mu}{\sigma_{11} + p \choose \sigma_{21}} - \NVRgrad u_{1} + {0 \choose \omega} &= \vect{0} & \text{ in } \Omega, \\
\frac{1}{2 \mu}{\sigma_{12} \choose \sigma_{22} + p} - \NVRgrad u_{2} - {\omega \choose 0} &= \vect{0}  & \text{ in } \Omega, \\
-\NVRdiv {\sigma_{11} \choose \sigma_{21}} &= f_{1} & \text{ in } \Omega, \\
-\NVRdiv {\sigma_{12} \choose \sigma_{22}} &= f_{2} & \text{ in } \Omega, \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\\
\vect{u} &= \vect{g_{D}} & \text{ on } \partial\Omega.
\end{align*}
Multiplying the first two equations by vector test functions $\vect{q}_{i}$ and the following three by scalar test functions $v_{i}$, and integrating by parts over an element $K$, we obtain
